### 2022-03-28 chapters/conclusions/conclusions.tex
commentaires patron

\begin{list}{}{%
    % \setlength{\listparindent}{-0.23in}%
    \setlength{\leftmargin}{0in}%
    % \setlength{\itemindent}{-0.23in} 
    }
    \item \textbf{1. How did the SCTLD spread through the FRT ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item The observed spread of the SCTLD through the FRT was successfully reproduced by modeling the transport of disease agents as neutrally buoyant material within the water column, and by calibrating a connectivity-based epidemiological model with field data. This therefore suggests that the disease spread through waterborne transmission. 
    \end{list}
    \item \textbf{2. What caused the apparent stalling of the propagation of SCTLD ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Our results suggest that there was an interruption of hydrodynamic-driven exchanges from the Florida Keys and the Marquesas to the DRTO during most of 2020, which prevented the propagation of disease agents westward. This interruption of the disease connectivity pathways to the DRTO was likely related to the local eddy activity generated by the meandering of the Florida Current.
    \end{list}
    \item \textbf{3. What was the impact of the PMDDP on the onset of the oubreak of SCTLD~?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Modeling of sediment transport suggests that the chopped rock particles produced by the non conventional dredging activities reached reefs where signs of disease were first reported during the first half of 2014. Furthermore, the modeled disease connectivity suggested that disease propagation from these sites to Virginia Key was possible prior to September 2014. The sediments released by the PMDDP might thus have triggered the onset of the SCTLD outreak at this site.
    \end{list}
    \item  \textbf{4. What is the effect of hurricane-induced wave-current interactions on transport processes and should they be taken into account when modeling the dispersal of coral larvae ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Results of a coupled wave-current model suggest that wave-induced transport under storm conditions, dominated by Stokes drift, can yield differences in particle trajectories of up to 20 km. Wave-current interactions should thus be taken into account to accurately model the dispersal of coral larvae during hurricanes.
    \end{list}
\end{list}

\modif{\section{Developments of the model}}

\modif{Answering the aforementioned research questions required the development of new modeling tools. As this thesis heavily relies on Lagrangian transport modeling, several features were added to the particle tracker of SLIM. Although the transport of the particles in itself was already implemented in the model, we developed features to characterize the state of the particles and compute connectivity indicators. The particle state can be used to represent biological traits such as the competency of coral larvae or abiotic characteristics such as the decaying mass of diseased material or the diameter of sediment particles. We developed tools to derive connectivity indicators such as the shortest paths between a pair sub-reef polygons in the obtained  connectivity network. Additionally, we developed an epidemiological model using the computed connectivity matrices to simulated the spread of the stony coral tissue loss disease throughout Florida. Finally, the particle tracker was made more flexible by adding the capability to transport particles using outputs from larger-scale ocean models such as HYCOM.}

\modif{Additionally, the coupling of SLIM with SWAN required some modifications of SLIM. First, we implemented a new parameterization of the wind drag coefficient to account for the formation of spume at the ocean-atmosphere interface in heavy wind conditions. Furthermore, a term was added to the model's equations to account for the gradient of atmospheric pressure generated at the center of the hurricane. Using SWAN also required the implementation of tools to process wave spectrum and parameters in SLIM. The implementation of the coupling between the two models was based on the ADCIRC+SWAN model. It required the development of tools to convert SLIM outputs and forcing into SWAN format, and to allow SLIM to read SWAN outputs. Compared to the ADCIRC+SWAN, an additional challenge that we had to overcome was the fact that SLIM is based on discontinous Galerkin finite element models while SWAN requires continuous fields on the mesh. This was especially critical when defining the communication between the nodes of the partitions of SWAN based on the partitioning of the mesh used in SLIM.}

@@

\modif{\subsection*{Better understanding the impact of hurricanes on corals}}
\modif{In this thesis, the impact of hurricanes on coral reef ecosystems was mostly evaluated by modeling wave-induced advective transport processes. However, when studying the propagation of drifting particles, assessing the impact of hurricanes on the dispersion of the particle cloud is also of interest. This could be performed by computing the distance between the positions of the particles of the cloud and its center of mass or by evaluating the evolution of the concentration of particles in the cloud. Furthermore, wave-induced transport is not the only way tropical cyclones affect coral reef biota. First, they cause mechanic damages to the reefs as they pass over them. However, this destruction can have positive consequences by facilitating asexual recruitment and cleaning algal overgrowth on the reefs \citep{banks2008reef}. Furthermore, hurricanes increase water turbidity and induce coral burial by resuspending sediments. These impacts could be assessed by combining the developed coupled wave-current model with our sediment transport model. Moreover, hurricane interact with the Florida Current by disrupting its vertical structure and reducing its transport. These structural perturbations might have consequences on eddy activity and baroclinic instabilities near the Florida Keys. Our study could therefore be extended by evaluating the impact of Hurricane Irma on the ocean state and the broader circulation patterns in Florida. Finally, by modifying the ocean circulation on their passage, hurricanes might influence the propagation of SCTLD through Florida's Coral Reef. Hurricane Irma could therefore have acted as a "super spreader" of disease agents or as a barrier to the propagation of the outbreak. This effect could be investigated by combining SLIM+SWAN with the developed epidemiological model.} \newline

>>>>

\begin{list}{}{%
    % \setlength{\listparindent}{-0.23in}%
    \setlength{\leftmargin}{0in}%
    % \setlength{\itemindent}{-0.23in} 
    }
    \item \textbf{1. How did the \modif{stony coral tissue loss disease} spread through the \modif{Florida Reef Tract} ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item The observed spread of the SCTLD through the FRT was successfully reproduced by modeling the transport of disease agents as neutrally buoyant material within the water column, and by calibrating a connectivity-based epidemiological model with field data. This therefore suggests that the disease spread through waterborne transmission. 
    \end{list}
    \item \textbf{2. What caused the apparent stalling of the propagation of \modif{disease} ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Our results suggest that there was an interruption of hydrodynamic-driven exchanges from the Florida Keys and the Marquesas to the DRTO during most of 2020, which prevented the propagation of disease agents westward. This interruption of the disease connectivity pathways to the DRTO was likely related to the local eddy activity generated by the meandering of the Florida Current.
    \end{list}
    \item \textbf{3. What was the impact of the \modif{PortMiami Deep Dredge Project} on the onset of the \modif{disease outbreak}?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Modeling of sediment transport suggests that the chopped rock particles produced by the non conventional dredging activities reached reefs where signs of disease were first reported during the first half of 2014. Furthermore, the modeled disease connectivity suggested that disease propagation from these sites to Virginia Key was possible prior to September 2014. The sediments released by the PMDDP might thus have triggered the onset of the SCTLD outreak at this site.
    \end{list}
    \item  \textbf{4. What is the effect of hurricane-induced wave-current interactions on transport processes and should they be taken into account when modeling the dispersal of coral larvae ?}
    \begin{list}{}{\setlength{\topsep}{0pt}}
        \item Results of a coupled wave-current model suggest that wave-induced transport under storm conditions, dominated by Stokes drift, can yield differences in particle trajectories of up to 20 km. Wave-current interactions should thus be taken into account to accurately model the dispersal of coral larvae during hurricanes.
    \end{list}
\end{list}

\modif{\section{Our contribution to the model development}}

\modif{Answering the aforementioned research questions required the development of new modeling tools. As this thesis relied heavily on Lagrangian transport modeling, several features were added to SLIM's particle tracker. Although the transport of the particles in itself was already implemented in the model, we developed features to characterize the state of the particles and compute connectivity indicators. The particle state can be used to represent biological traits such as the competency of coral larvae or abiotic characteristics such as the decay rate of diseased agent mass or the diameter of sediment particles. Post-processing features were developed to compute the concentration or the deposition of particles. We also developed tools to derive connectivity indicators such as the shortest paths between a pair sub-reef polygons in the connectivity networks. Additionally, we developed an epidemiological model using the computed connectivity matrices to simulated the spread of the stony coral tissue loss disease throughout Florida's Coral Reef. Finally, the particle tracker was made more flexible by developing new readers for different ocean model outputs such as HYCOM or the Nucleus for European Modelling of the Ocean (NEMO; \citealp{madec2017nemo}).}

\modif{Additionally, coupling SLIM with SWAN required some modifications of SLIM. First, we implemented a new parameterization of the wind drag coefficient to account for the formation of spume at the ocean-atmosphere interface in heavy wind conditions. An additional term was also added to the model's equations to account for the gradient of atmospheric pressure generated at the center of the hurricane. Using SWAN also required the implementation of tools to process wave spectrum and parameters in SLIM. The implementation of the coupling between the two models followed the same approach as the ADCIRC+SWAN model. It required the development of tools to convert SLIM outputs and forcing into SWAN format, and to allow SLIM to read SWAN outputs. Compared to ADCIRC+SWAN, an additional challenge that we had to overcome was the fact that SLIM is based on discontinous Galerkin finite element models while SWAN requires continuous fields on the mesh. This was especially critical when defining the communication between the nodes of the partitions of SWAN based on the partitioning of the mesh used in SLIM.}

@@

\modif{\subsection*{Better understanding the impact of hurricanes on corals}}
\modif{In this thesis, the impact of hurricanes on coral reef ecosystems was mostly evaluated by modeling wave-induced advective transport processes. However, when studying the propagation of drifting particles, assessing the impact of hurricanes on the dispersion of the particle cloud is also of interest. This could be performed by computing the distance between the positions of the particles of the cloud and its center of mass or by evaluating the evolution of the concentration of particles in the cloud. Furthermore, wave-induced transport is not the only way tropical cyclones affect coral reef biota. First, they can cause a wholesale destruction of the reefs as they pass over them. However, this destruction can have positive consequences by facilitating asexual recruitment and cleaning algal overgrowth on the reefs \citep{banks2008reef}. Furthermore, hurricanes increase water turbidity and induce coral burial by resuspending sediments. These impacts could be assessed by combining the developed coupled wave-current model with our sediment transport model. Moreover, hurricane interact with the Florida Current by disrupting its vertical structure and reducing its transport. These structural perturbations might have consequences on eddy activity and baroclinic instabilities near the Florida Keys. Our study could therefore be extended by evaluating the impact of Hurricane Irma on the ocean state and the broader circulation patterns in Florida although baroclinic models might be required. Finally, by modifying the ocean circulation on their passage, hurricanes might influence the connectivity of the Florida Reef Tract. The coupled SLIM+SWAN model could thus be used to evaluate the impact of Hurricane Irma on coral larval connectivity as a continuation of the work initiated in chapter \ref{chap:irma}. Furthermore, the hurricane-induced transport might have impacted the propagation of SCTLD through Florida's Coral Reef. Hurricane Irma could therefore have acted as a "super spreader" of disease agents or as a barrier to the propagation of the outbreak. This effect could be investigated by combining SLIM+SWAN with the developed epidemiological model.} \newline

### 2022-03-28 chapters/intro/intro.tex
commentaires patron

In this thesis, we used the 2D barotropic version of the unstructured-mesh Second-generation Louvain-la-Neuve Ice-ocean Model (\href{https://www.slim-ocean.be/}{SLIM}). This model relies on Discontinuous Galerkin finite element methods \citep{aizinger2002discontinuous} and has already been successfully used to model the hydrodynamics of the whole Great Barrier Reef \citep{lambrechts2008multi}. \modif{The 2D version of SLIM solves the shallow water equations, which assume a constant water density and compute depth-averaged current velocities. This model is therefore well suited for shallow areas where the water-column is well mixed. Furthermore, assuming homogeneous vertical profiles reduces the computational cost and allows a finer spatial resolution of the model. In deeper areas however, 3D baroclinic models are required to capture the variations of density in the water column due to pressure, temperature and salinity. These differences in density cause baroclinic instabilities that generate mesoscale eddies. However, 3D baroclinic models are computationally more expensive, which prohibits the use of a reef-scale resolution on the whole Florida Reef Tract. As most reefs of the Florida are located in areas shallower than 30 meters (Fig. \ref{intro:bathy}), we coupled the 2D version of SLIM with GOM-HYCOM to indirectly capture the impact of baroclinic instabilities on the connectivity of the Florida Keys. Using this approach, \cite{frys2020fine} successfully modeled coral larval connectivity in the FRT, \ie~the exchanges of larvae between reefs. The results of the model used in this thesis were validated against observations and showed good agreement with tide gauge and ADCP measurements, as illustrated in appendix \ref{onset:appendice} and chapter \ref{chap:irma}}.

@@

\modif{Transport processes are at the core of this thesis. They can be estimated using two general methods: Eulerian tracers and Lagrangian particles. Eulerian methods were the earliest methods applied for transport modeling and are still widely used today. They offer the advantages of a fixed grid by solving the transport equations directly on the native grid of the model and are easy to implement However, mass conservation issues may arise in problems with high initial gradient of concentrations \citep{spivakovskaya2007lagrangian}. The Lagrangian methods, on the other hand, follow virtual particles of zero spatial extent whose movements are modeled with a stochastic differential equation consistent with the advection-diffusion equation \citep{spivakovskaya2007lagrangian,van2018lagrangian}. One key difference between the two  methods is their computational cost. For each time step, the movements of $N$ virtual particles require $N$ sets of computation, while the advection-diffusion of a tracer takes as many sets of computation as there are discrete grid cells in the model. Additionally, Lagrangian methods allow the storage of the entire trajectory history of the virtual particles. This enables the \textit{a posteriori} computation of connectivity between different regions of the ocean \citep{thomas2014numerical,frys2020fine}. Furthermore, particles can be advected in offline mode when velocity fields are stored in order to reduce computation time, while Eulerian tracers' equations must be solved online with the current velocities. Finally, Lagrangian particles can be advected backward in time. This reverse-time analysis gives a spatial probability distribution of places from where it is likely that particles originated to reach a given location \citep{spivakovskaya2005simulation}. Transport processes were therefore studied in this thesis using Lagrangian particles advected by SLIM currents. To answer our different research questions, we adapted SLIM transport model to evaluate the dispersal of the causative agent of the SCTLD and sediments particles.}

@@

\section*{Outline}
This thesis is mainly a collection of publications describing our answers to our four main research questions. Chapter \ref{chap:sctld} describes the development of a coupled epidemio-hydrodynamic model to reproduce the observed spread of SCTLD in the FRT. Chapter \ref{chap:drto} explains how this model was used to study the apparent stalling of the spread of the SCTLD between the Marquesas and the DRTO. Chapter \ref{chap:onset} investigates the impact of the PMDDP on the onset of the SCTLD oubreak in 2014. Chapter \ref{chap:irma} describes the development of a coupled wave-current model to study the impacts of Hurricane Irma (2017) on wave-induced transport processes in the Florida Keys. \modif{By lack of time, the propagation of coral larvae during Irma was not studied in this chapter}. Finally, chapter \ref{chap:conclusion} presents the conclusions of this thesis and perspectives for further works.

>>>>

In this thesis, we used the 2D barotropic version of the unstructured-mesh Second-generation Louvain-la-Neuve Ice-ocean Model (\href{https://www.slim-ocean.be/}{SLIM}). This model relies on Discontinuous Galerkin finite element methods \citep{aizinger2002discontinuous} and has already been successfully used to model the hydrodynamics of the whole Great Barrier Reef [ADD OTHER STUFF] \citep{lambrechts2008multi}. \modif{The 2D barotropic version of SLIM solves the shallow water equations, which assume a constant water density and compute depth-averaged current velocities. This model is therefore well suited for shallow areas where the water column is well mixed. Furthermore, assuming homogeneous vertical profiles reduces the computational cost and allows a finer spatial resolution of the model. However, in deeper areas or in regions where there is an inflow of freshwater, 3D baroclinic models are required to capture the variations of density in the water column due to temperature and salinity gradients. These differences in density can cause baroclinic instabilities that generate mesoscale eddies. Nonetheless, 3D baroclinic models are computationally more expensive, which prohibits the use of a reef-scale resolution on the whole Florida Reef Tract. As most Florida reefs are located in areas shallower than 30 meters (Fig. \ref{intro:bathy}), we coupled the 2D version of SLIM with GOM-HYCOM to indirectly capture the impact of baroclinic instabilities on the connectivity of the Florida Keys. Using this approach, \cite{frys2020fine} successfully modeled coral larval connectivity in the FRT with SLIM, \ie~the exchanges of larvae between reefs. [TODO: move somewhere else]The results of the model used in this thesis were validated against observations and showed good agreement with tide gauge and ADCP measurements, as illustrated in appendix \ref{onset:appendice} and chapter \ref{chap:irma}}.

@@

\modif{Transport processes are at the core of this thesis. They can be estimated using two general methods: Eulerian tracers and Lagrangian particles. Eulerian methods were the earliest methods applied for transport modeling and are still widely used today. They offer the advantages of a fixed grid by solving the transport equations directly on the native grid of the model and are easy to implement \citep{spivakovskaya2007lagrangian}. The Lagrangian methods, on the other hand, follow virtual particles whose movements are modeled with a stochastic differential equation consistent with the advection-diffusion equation \citep{spivakovskaya2007lagrangian,van2018lagrangian}. One key difference between the two  methods is their computational cost. For each time step, the movements of $N$ virtual particles require $N$ sets of computation, while the advection-diffusion of a tracer takes as many sets of computation as there are discrete grid cells in the model. Additionally, Lagrangian methods allow the storage of the entire trajectory history of the virtual particles. This enables the \textit{a posteriori} computation of connectivity between different regions of the ocean \citep{thomas2014numerical,frys2020fine}. Finally, Lagrangian particles can be advected backward in time in offline mode, when velocity fields are stored. This reverse-time analysis gives a spatial probability distribution of places from where it is likely that particles originated to reach a given location \citep{spivakovskaya2005simulation}. In this thesis, transport processes were therefore studied using Lagrangian particles advected by SLIM currents. To answer our different research questions, we adapted SLIM transport model to evaluate the dispersal of the causative agent of the SCTLD and sediments particles.}

@@

\section*{Outline}
This thesis is mainly a collection of publications describing our answers to our four main research questions. Chapter \ref{chap:sctld} describes the development of a coupled epidemio-hydrodynamic model to reproduce the observed spread of SCTLD in the FRT. Chapter \ref{chap:drto} explains how this model was used to study the apparent stalling of the spread of the SCTLD between the Marquesas and the DRTO. Chapter \ref{chap:onset} investigates the impact of the PMDDP on the onset of the SCTLD oubreak in 2014. Chapter \ref{chap:irma} describes the development of a coupled wave-current model to study the impacts of Hurricane Irma (2017) on wave-induced transport processes in the Florida Keys. Finally, chapter \ref{chap:conclusion} presents the conclusions of this thesis and perspectives for further works.

### 2022-03-28 chapters/irma/article.tex
commentaires patron

We followed the approach of \cite{dietrich2012performance} by characterizing the wave-induced forces on currents using the radiation-stress (RS) gradient formalism, which has been successfully applied in both 2D and 3D coupled wave-current models under storm conditions \citep{hope2013hindcast, sebastian2014characterizing, brown2013depth}. \modif{In this formalism, wave-averaged effects are characterized by the divergence of a stress tensor representing the flux of momentum to currents due to surface gravity waves \citep{longuet1964radiation,lane2007wave,kumar2012implementation}}. An alternative formalism is the vortex-force (VF) representation \citep{mcwilliams2004asymptotic}, that provides a clearer decomposition of the wave effect \citep{lane2007wave}. \modif{In this formalism, wave impact on current is represented by the combination of the gradient of a Bernoulli head representing adjustments to the pressure in incompressible fluids, and the interaction of the Stokes drift with the vorticity of the flow, \textit{i.e.} the vortex force \citep{lane2007wave}}. Although both approaches were adopted by coastal modeling communities, there is an ongoing scientific debate over the correctness and applicability of the two concepts \citep{ardhuin2008comments, mellor2013waves, mellor2015combined, ardhuin2017comments}. \cite{xia2020implementation} recently implemented the two formalisms in a 3D unstructured-grid model and compared them in three typical coastal systems, showing that the 3D RS algorithm could generate unrealistic offshore currents near shorelines. Despite these shortcomings, the 3D RS method reproduced most wave-induced currents and the 2D RS formalism remains a well-validated modeling approach. Furthermore, \cite{mellor2013waves} showed that the RS approach was valid when $[\frac{\partial H}{\partial\mathbf{x}}/\sinh(kh)]^2$ is smaller or of the same order as $(ka)^2$, where $a$ is the wave amplitude. We evaluated these quantities and verified that the validity criterion was met in our model domain. Additionally, since the VF and RS approached are formally equivalent \citep{lane2007wave}, we selected the 2D RS formalism, as it has the advantage of summarizing the impact of waves on the currents in a single additional stress term in the hydrodynamic model equations.

>>>>

We followed the approach of \cite{dietrich2012performance} by characterizing the wave-induced forces on currents using the radiation-stress (RS) gradient formalism, which has been successfully applied in both 2D and 3D coupled wave-current models under storm conditions \citep{hope2013hindcast, sebastian2014characterizing, brown2013depth}. \modif{In this formalism, wave-averaged effects are characterized by the divergence of a stress tensor representing the flux of momentum to currents due to surface gravity waves \citep{longuet1964radiation,lane2007wave,kumar2012implementation}}. An alternative formalism is the vortex-force (VF) representation \citep{mcwilliams2004asymptotic}, that provides a clearer decomposition of the wave effect \citep{lane2007wave}. \modif{Here, wave impact on current is represented by the combination of the gradient of a Bernoulli head representing adjustments to the pressure in incompressible fluids, and the interaction of the Stokes drift with the vorticity of the flow, \textit{i.e.} the vortex force \citep{lane2007wave}}. Although both approaches were adopted by coastal modeling communities, there is an ongoing scientific debate over the correctness and applicability of the two concepts \citep{ardhuin2008comments, mellor2013waves, mellor2015combined, ardhuin2017comments}. \cite{xia2020implementation} recently implemented the two formalisms in a 3D unstructured-grid model and compared them in three typical coastal systems, showing that the 3D RS algorithm could generate unrealistic offshore currents near shorelines. Despite these shortcomings, the 3D RS method reproduced most wave-induced currents and the 2D RS formalism remains a well-validated modeling approach. Furthermore, \cite{mellor2013waves} showed that the RS approach was valid when $[\frac{\partial H}{\partial\mathbf{x}}/\sinh(kh)]^2$ is smaller or of the same order as $(ka)^2$, where $a$ is the wave amplitude. We evaluated these quantities and verified that the validity criterion was met in our model domain. Additionally, since the VF and RS approached are formally equivalent \citep{lane2007wave}, we selected the 2D RS formalism, as it has the advantage of summarizing the impact of waves on the currents in a single additional stress term in the hydrodynamic model equations.

### 2022-03-28 chapters/onset/article.tex
commentaires patron

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapters/onset/figures/fig_mesh_onset.png}
	\caption{\textbf{A}: Model mesh near the dredged channel. Elements have a characteristic length of 100 m over reefs (in light grey) and along the coasts (in dark gray). The monitoring sites considered in the present study are shown by blue dots. The date of the first reported signs of SCTLD at these sites is given between brackets. The Ocean Dredge Material Disposal Site (ODMDS) is shown in red and the discharge pipe of the Miami Central District Municipal Wastewater Treatment Plant in green. \textbf{B}: Close up view of the dredged channel. The locations of the different types of dredging that took place during the expansion of PoM are shown by colored markers. \textbf{C:} Snapshot of the modeled currents in the vicinity of the dredged channel. Small-scale flow features such as the acceleration of currents between reefs and islands are well reproduced by the model.}
	\label{fig:onset_mesh}
\end{figure}

@@

The potential impact of dredging on the onset of the SLD outbreak was more specifically assessed at five monitoring sites were signs of disease were reported. Four sites were reefs where disease was reported in 2014 by \cite{precht2016unprecedented}. The fifth site was the monitoring station HBSC1-CP, used throughout the expansion of PoM, where signs of disease were already reported in May 2014 (Fig. \ref{fig:onset_mesh}A, see supplementary material \ref{onset:appendice}). The impact of dredging was assessed by counting the total number of sediment particles originating from each dredging that were transported within 500 m of all five site. This distance was chosen to match the scale of the 500 m $\times$ 500 m reef polygons used in the model \citep{dobbelaere2020coupled}. The resulting numbers of particles were then divided by the total number of sediment particles released by each type of dredging operation. Larger values of this indicator would suggest a greater impact of a given type of dredging operation at a given monitoring sites.

@@

\begin{subappendices}
	\section{Reported disease signs at site HBSC1-CP in May 2014}\label{onset:appendice}
	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{chapters/onset/figures/hbsc1_cp.png}
		\caption{Pictures of a monitored colony at site HBSC1-CP. The first signs of the disease appeared between May 25 and May 30. By June 4, the whole colony was completely dead}
	\end{figure}
\end{subappendices}

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapters/onset/figures/fig_mesh_onset.png}
	\caption{\textbf{A}: Model mesh near the dredged channel. Elements have a characteristic length of 100 m over reefs (in light grey) and along the coasts (in dark gray). The monitoring sites considered in the present study are shown by blue dots. The date of the first reported signs of SCTLD at these sites is given between brackets. The Ocean Dredge Material Disposal Site (ODMDS) is shown in red and the discharge pipe of the Miami Central District Municipal Wastewater Treatment Plant in green. \textbf{B}: Close up view of the dredged channel. The locations of the different types of dredging that took place during the expansion of PortMiami are shown by colored markers. \textbf{C:} Snapshot of the modeled currents in the vicinity of the dredged channel. Small-scale flow features such as the acceleration of currents between reefs and islands are well reproduced by the model.}
	\label{fig:onset_mesh}
\end{figure}

@@

The potential impact of dredging on the onset of the SLD outbreak was more specifically assessed at five monitoring sites were signs of disease were reported. Four sites were reefs where disease was reported in 2014 by \cite{precht2016unprecedented}. The fifth site was the monitoring station HBSC1-CP, used throughout the expansion of PoM, where signs of disease were already reported in May 2014 (\citealp{dial2015}; Fig. \ref{fig:onset_mesh}A, see supplementary material \ref{onset:appendice}). The impact of dredging was assessed by counting the total number of sediment particles originating from each dredging that were transported within 500 m of all five site. This distance was chosen to match the scale of the 500 m $\times$ 500 m reef polygons used in the model \citep{dobbelaere2020coupled}. The resulting numbers of particles were then divided by the total number of sediment particles released by each type of dredging operation. Larger values of this indicator would suggest a greater impact of a given type of dredging operation at a given monitoring sites.

@@

\begin{subappendices}
	\section{Reported disease signs at site HBSC1-CP in May 2014}\label{onset:appendice}
	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{chapters/onset/figures/hbsc1_cp.png}
		\caption{Pictures of a monitored colony at site HBSC1-CP \citep{dial2015}. The first signs of the disease appeared between May 25 and May 30. By June 4, the whole colony was completely dead}
	\end{figure}
\end{subappendices}

### 2022-03-28 chapters/sctld/article.tex
commentaires patron

The spread of the SCTLD throughout the FRT was simulated using a connectivity-based \cite{kermack1927contribution} SIR model. SIR models are among the most standard epidemiological models. They divide individuals into three compartments: susceptible ($S$), infectious ($I$) and removed ($R$). When affected by the disease, susceptible individuals become infectious and infect other susceptible individuals until they are removed from the system, either through recovery or death. Such models usually rely on the hypothesis of an homogeneous, well-mixed population. To account for the spatial heterogeneity of the FRT, the basic SIR formulation was modified by considering the fractions of susceptible ($S_j$), infectious ($I_j$) and removed ($R_j$) corals of each sub-reef $j$. Although the pathogen of SCTLD has not been identified, studies suggest a contagious mode of transmission \citep{aeby2019pathogenesis,muller2020spatial}. Here, we use 'infectious' to denote disease agents that could be passed from individual to individual, which are responsible for causing disease signs. We note that the disease agents could be biological or non-biological in nature. In the present study's epidemiological model, individual reefs interact through the exchange of disease agents as represented by the different connectivity matrices. For each sub-reef $j$ and at any time, the following relations hold: $0\leq S_j,I_j,R_j\leq 1$ and $S_j+I_j+R_j=1$. The evolution of these fractions through time is governed by the following equations:

>>>>

The spread of the SCTLD throughout the FRT was simulated using a connectivity-based \cite{kermack1927contribution} SIR model. SIR models are among the most standard epidemiological models. They divide individuals into three compartments: susceptible ($S$), infectious ($I$) and removed ($R$). When affected by the disease, susceptible individuals become infectious and infect other susceptible individuals until they are removed from the system, either through recovery or death. \modif{Other more advanced compartmental models take reinfection into account (SIS) or consider a latent non-infectious exposed ($E$) state between the susceptible and infectious states (SEIR) \citep{Murray07}. However, due to the high mortality of the disease and in the absence of clear evidence suggesting a latency period following infection, the SIR model seemed a reasonably simple and generic modeling choice for this study}. Such models usually rely on the hypothesis of an homogeneous, well-mixed population. To account for the spatial heterogeneity of the FRT, the basic SIR formulation was modified by considering the fractions of susceptible ($S_j$), infectious ($I_j$) and removed ($R_j$) corals of each sub-reef $j$. Although the pathogen of SCTLD has not been identified, studies suggest a contagious mode of transmission \citep{aeby2019pathogenesis,muller2020spatial}. Here, we use 'infectious' to denote disease agents that could be passed from individual to individual, which are responsible for causing disease signs. We note that the disease agents could be biological or non-biological in nature. In the present study's epidemiological model, individual reefs interact through the exchange of disease agents as represented by the different connectivity matrices. For each sub-reef $j$ and at any time, the following relations hold: $0\leq S_j,I_j,R_j\leq 1$ and $S_j+I_j+R_j=1$. The evolution of these fractions through time is governed by the following equations:
